The effects of linkage disequilibrium on maintaining balanced polymorphisms in seasonal adaptation

### 2019-12-02 writing/manuscript.Rmd

# Abstract

# Introduction
Something about an orchard in Pennsylvania [@Bergland2014]. As an example of how to cite.

>>>>

# Abstract
Understanding the mechanisms of preserving genetic diversity is paramount to predicting the capacity of organisms to adapt to changing environmental conditions. In environments characterized by seasonal fluctuations, certain alleles may be beneficial under certain conditions and harmful in other environments. This fluctuation of additive benefits results in a strong directional selection that varies according to the season. @Bergland2014 experimentally observed that seasonal variation is able to preserve polymorphisms that maintain genetic diversity. @Wittman2017 modeled how allelic dominance changes according to the season in an organism to account for different seasonal life-history strategies. They present a term called “segregation lift” which describes instances of seasonal allele-fluctuations which serve to maintain genetic diversity over time. In this paper, we build upon the @Wittman2017 model, using the same parameters to study how genetic linkage and seasonal balance may affect segregation lift. The @Wittman2017 model found that in an unlinked genome the dominance parameter needed to be greater than 0.5 for segregation lift to occur. Our model suggests that in diploid organisms exhibiting genetic linkage, segregation lift occured when dominance=0.5, suggesting that multilocus polymorphisms are maintained if, at a certain locus, one allele over the other  is beneficial during one season.

# Introduction

Quantitative genetic theory suggests that plasticity occurs in response to selective pressure exerted by environmental heterogeneity [@Cooper2014].  Phenotypic plasticity is especially beneficial in variable environments because genetic variability allows for adaptation to changing conditions [@Gulisija2016]. Fluctuating selection can maintain polymorphisms in seasonal environments if strong selective pressure towards both homozygous genotypes varies seasonally [@Novak2017, @Bergland2014]. Understanding the role of seasonality on adaptation can allow more accurate predictions to be made about the capacity for organisms to adapt to climate change conditions [@Fournier-Level2016].

There has been some debate regarding whether balancing selection is capable of maintaining genetic polymorphisms in highly variable environments, and it is often thought that  this mechanism alone is not enough to maintain balanced polymorphism. (@Asthana2005, @Hedrick). However a study by @Bergland2014 observing Drosophila melanogaster in a temperate Pennsyvanian orchard revealed that strong seasonal selection can drive balancing selection favoring different life history traits over a short time span. These results suggest that seasonal variation contributes to the maintenance of two alleles to be present at the same loci long-term.  If there is no fitness cost associated with polymorphic variation and there is a benefit gained through heterozygosity, then it is suggested that that balancing selection could preserve genetic variation [@Asthana2005, @Wittman2017].

@Wittman2017 created a model simulating the seasonal selection observed by @Bergland2014 to clarify the role this allelic fluctuation plays in maintaining multilocus polymorphisms. This model introduces the term “segregation lift” which considers the importance of dominance in maintaining polymorphism across many loci in the genome and proposes that segregation lift plays an important role in preserving genetic variation.

Our model recreates and builds upon the work of @Wittman2017 to understand how linkage disequilibrium contributes to maintaining polymorphisms in a natural population. We study how rate of crossover, epistasis, allelic dominance, and seasonal balance contributes to segregation lift.
